---
title: 'Major Project Inventory: data quality'
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: https://github.com/bcgov/mpi_data_quality
    css: style.css
runtime: shiny
resource_files:
- processed_data/plots.rds
- processed_data/mpi_raw.rds
---

```{r global, include=FALSE}
# Copyright 2022 Province of British Columbia
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and limitations under the License.
# libraries----------
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
tbbl <- read_rds(here("processed_data", "mpi_raw.rds"))|>
  select(project_id, project_name, estimated_cost, project_stage, project_status, last_update)|>
  mutate(project_stage=if_else(project_stage=="Elementary School Additions", NA, project_stage))

pp_name <- tbbl|>
  filter(
    (project_status=="Proposed" & is.na(project_stage))|
    (project_status=="Construction started" & project_stage=="Tender/Preconstruction")
    )|>
  select(project_name, estimated_cost)|>
  distinct()

pp <- semi_join(tbbl, pp_name|>select(project_name))|>
  arrange(project_name, last_update)|>
  group_by(project_name)|>
  pivot_longer(cols=c(project_stage, project_status))

projects <- pp_name|>
  arrange(desc(estimated_cost))|>
  pull(project_name)

plots <- readRDS(here::here("processed_data", "plots.rds")) %>%
  mutate(facet_by = str_to_title(str_replace_all(facet_by, "_", " ")))%>%
  filter(facet_by!="Project Type")
```

Raw vs Clean
======================================================


Inputs {.sidebar data-width=400}
-------------------------------------

```{r}
selectInput(
  "facet_var",
  "What variable do you want to facet by?",
  plots$facet_by,
  selected = plots$facet_by[1],
  multiple = FALSE
)
```

* Raw data is sourced from [Economic Development](https://www2.gov.bc.ca/gov/content/employment-business/economic-development/industry/bc-major-projects-inventory/recent-reports) with no processing.

* Clean data corrects for two main problems in the data:
  + Mis-categorization: some projects change categories mid-project. 
  + Missing values, both explicit (the presence of an absence i.e. value recorded as NA) and implicit (the absence of a presence i.e. project missing from MPI).

* In order to clean the data:
  + For each project the reported categories are replaced with that project's modal (most common) reported category.
  + Implicit missing values are converted to explicit missing values by left joining 1) a regular grid of quarterly dates with 2) the observed data.
  + Finally the explicit missing variables (estimated_cost, telephone, project_status, first_entry_date) are "updown" filled:  First missing values are filled up (backwards in time) and then down (forwards in time).

* Note the variable `project_stage` is not cleaned as it is possible that NA is a valid stage, not a missing value: see problem and solution pages.

Column
-----------------------------------------------------------------------

###

```{r, fig.retina=2}
renderPlot({
  req(input$facet_var)
  plt <- plots %>%
    filter(facet_by == input$facet_var) %>%
    select(plot) %>%
    pull()
  plt[[1]]
}) %>%
  bindCache(input$facet_var)
```

Problem
===================================================================


Inputs {.sidebar}
-------------------------------------

-   In the table to the right, the red cell entry (status=Proposed, stage=NA) is potentially problematic: 
    - is NA a valid level of project stage for proposed projects?
    - or does NA indicate that project stage is one of Consultation/Approvals, Permitting, Preliminary/Feasibility, Tender/Preconstruction, but we do not know which one?


```{r}
selectInput(
  "project_id",
  "Select a project from the red cell:",
  projects,
  projects[1]
)

filtered <- reactive({
  pp|>
    filter(project_name==input$project_id)
})
```
 
Column
-------------------------------------

### Tabulation of project stage and project status

```{r}
tabled <- with(tbbl, table(project_stage, project_status, useNA="ifany"))
cells_to_format <- list(c(5, 4)) 
for (cell in cells_to_format) {
  row <- cell[1]
  col <- cell[2]
  tabled[row, col] <- cell_spec(
    tabled[row, col],
    format = "html", # Required for HTML styling
    color = "red",
    bold = TRUE
  )
}
kable(tabled, "html", escape = FALSE)|>
  kable_styling()
```

### Plot of stage and status
    
```{r}
renderPlot({
  filtered()|>
    ggplot(aes(last_update, value, colour=name, size=estimated_cost))+
    geom_point()+
    labs(title=filtered()$project_name[1])
})
```

Solution
=================================================

-   I suggest `project_stage` and `project_status` get collapsed into a single variable `project_stagus`, with mutually exclusive, totally exhaustive levels:
    
    - Completed
    - Construction started	
    - On hold
    - Proposed (for projects prior to any of the levels below)
    - Consultation/Approvals
    - Permitting
    - Preliminary/Feasibility
    - Tender/Preconstruction
    
    
-   If `project_stagus` is either blank or NA the meaning is clear: the `project_stagus` was not available this quarter.
  



